---
title: "Filter/joins (like VLOOKUP in Excel)"
author: "Sarah Antonelli"
date: "12/8/2019"
output: html_document
---

## Attach packages

tidyverse, readxl, here
```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Read in 3 files:

- invert_counts_curated.xlsx
- fish_counts_curated.csv
- kelp_counts_curated.xlsx

```{r}
invertCounts <- read_excel(here("R-workshop-data", "invert_counts_curated.xlsx"))

fishCounts <-read_csv(here("R-workshop-data", "fish_counts_curated.csv"))

kelpCountsAbur <- read_excel(here("R-workshop-data", "kelp_counts_curated.xlsx"))
```

- select() function: keep, exclude, or rearrange COLUMNS
- filter() function: keep or exclude ROWS (observations) based on whether or not they match variable conditions

## Filter to match rows based on a character string

Only keep observations from fishCounts where the common_name variable matches "garibaldi"
```{r}

fishGaribaldi <- fishCounts %>%
  filter(common_name == "garibaldi")

# Use janitor::clean_names() to automatically update all of your column names to lowercase_snake_case as default. You can also set it to camelCase or anything else
```

## Filter to keep or exclude observations based on numeric conditions

Create a subset where total_counts is greater than or equal to 50:

```{r}
fishOver50 <- fishCounts %>%
  filter(total_count >= 50)
```

Create a subset that only contains observations for fish at Mohawk Reef ("mohk"):

```{r}
mohawkFish <- fishCounts %>% 
  filter(site == "mohk")
```

## Filter to keep rows that match THIS OR THAT

Example: We want to keep any observations for garibaldi, blacksmith, or black surfperch

Use the %in% operator to look for any matches within a vector
```{r}
fish3SP <- fishCounts %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))

# Writing OR statements explicitly:
# Use the vertical line | to indicate "or"

fish3SPv2 <- fishCounts %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith" |
           common_name == "black surfperch" )
```

## Filter to return rows that match conditions for MULTIPLE variables (AND statements)

I want to create a subset called urchinAburMohk that only keeps observations where:
- site is "abur" or "mohk" AND
- species is "purple urchin"

Some ways to do this:
- separate "AND" conditions within a single filter line by a comma
- separate "AND" conditions within a single filter line by an ampersand (&)
- putting each filter condition within a separate filter line

```{r}
urchinAburMohk <- invertCounts %>% 
  filter(site %in% c("abur","mohk"), common_name == "purple urchin")

# EXAMPLE PIPE ORDER:
# pipeTest <- fishCounts %>% 
#   select(common_name, total_count) %>% 
#   filter(year == 2016)
```


Activity: create a subset from fishCounts, called lowGaribaldiRw that only contains:

- observations for "garibaldi" and "rock wrasse"
- where the total_count is LESS than 10

```{r}
lowGaribaldiRw <- fishCounts %>% 
  filter(common_name %in% c("garibaldi", "rock wrasse"),
         total_count < 10)
```


## Filter rows that DO NOT MATCH conditions

Use != (indicates "does NOT match")

Try to create a subset from fishCounts that does NOT contain observations for Mohawk Reef ("mohk"):

```{r}
noMohawkReef <- fishCounts %>% 
  filter(site != "mohk")

# You can use str_match to search for partial matches. Would need to set the pattern appropriately
```

An example of filter as part of a piped sequence:

```{r}
senoritaAllYears <- fishCounts %>% 
  filter(common_name == "senorita") %>% 
  group_by(site) %>% 
  summarize(countAllYears = sum(total_count))
```

## A couple ways of joining data frames

- `full_join()`: keeps everything (safe join)
- `left_join()`: if we are merging A and B, will only retain observations from B that have matches in A

```{r}
fish2016_2017 <- fishCounts %>% 
  filter(year == 2016 | year == 2017)
```

Join that data frame with kelpCountsAbur
```{r}
# aburKelpJoin <- fish2016_2017 %>%
#   full_join(kelpCountsAbur, by = c("year", "site"))

# This failed at first because one data frame had year as text and one had year as a double
```

```{r}
kelpCountsAbur <- kelpCountsAbur %>% 
  mutate(year = as.numeric(year))
```

Now we can join them: 
```{r}
aburKelpJoin <- fish2016_2017 %>%
  full_join(kelpCountsAbur, by = c("year", "site"))

ggplot(aburKelpJoin, aes(x = year, y = total_count)) +
  geom_col(aes(fill = site), position = "dodge")

```