---
title: "Intro to readxl"
author: "Amanda and Sarah Antonelli"
date: "12/7/2019"
output: html_document
---

# Learning `readxl`

We are working with data in readxl and it's going to be amazing!

We're going to use the following packages:

- tidyverse
- readxl
- writexl
- here

```{r packages}
library(tidyverse)
library(readxl)
library(writexl)
library(here)

# install.packages("packageName")
```

### Read in fish counts

```{r read in fish counts data}
##### Read in the fish_counts_curated.csv
fishCounts <- read_csv(here("R-workshop-data","fish_counts_curated.csv"))

# fishCounts <- read_csv("R-workshop-data/fish_counts_curated.csv")
```

### Read in data from a single Excel worksheet

Now, to read in an excel spreadsheet, we'll use the function `read_excel`.


```{r read in substrate data}
substrateCover <- read_excel(here("R-workshop-data","substrate_cover_curated.xlsx"),
                             skip = 4,
                             na = "-9999")
# if you have multiple na indicators in your raw data, you can use a combine function like. Ex: na = c("-9999","-999")
```

### Use read_excel() to read in a range of values

```{r read in part of substrate data}
carpCover2000 <- read_excel(here("R-workshop-data","substrate_cover_curated.xlsx"),
                            range = "D12:I15",
                            col_names = c("site", "transect", "quad", "side", "type", "coverage"))
```

### How can we read in data from an Excel workbook with MULTIPLE worksheets?





