---
title: "Intro to ggplot2"
author: "Sarah Antonelli"
date: "12/8/2019"
output: html_document
---

## Attach packages

```{r packages}
library(tidyverse)
library(readxl)
library(here)
```

## Read in the data

```{r name data objects}
caNp <- read_csv(here("R-workshop-data", "ca_np.csv"))
ciNp <- read_excel(here("R-workshop-data", "ci_np.xlsx"))
```

## Make a graph

1. Tell R we're using `ggplot`
2. Tell R what data we're using and what variable we want to plot
3. Tell R what type of graph we're creating (ie geom)

```{r graph structure, eval = FALSE}
ggplot(data = THIS, aes(x = THIS, y = THIS)) +
  geom_THIS()
```

Now let's create a graph of CI visitation:
```{r create CI NP line graph}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_line()
```

```{r create CI NP scatterplot}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_point()
```

1. Create a column graph with these data
2. Create an area graph with these data
```{r column graph}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_col()
```

```{r area graph}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_area()
```

## Intro to customizing `ggplot` graphs

- color = update point or line color
- fill = update the fill characteristic of an element
- linetype = 
- pch = value from 1 to 30 for point type
- size = change size of elements
- alpha = change transparency of an element

```{r customize plot with ggplot}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_line(color = "limegreen",
            linetype = "dashed")
```

```{r customize plot with ggplot 2}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_point(color = "blue",
            pch = 17,
            size = 4,
            alpha = 0.5)
```

I have some awesome text here, and below is my amazing finalized graph:

```{r, echo = FALSE}
ggplot(data = ciNp, aes(x = year, y = visitors)) +
  geom_point(color = "royalblue4",
            pch = 21,
            size = 4,
            alpha = 0.8)
```