---
title: "Pivot Tables"
author: "Sarah Antonelli"
date: "12/8/2019"
output: html_document
---

## Attach libraries

```{r}

library(tidyverse)
library(readxl)
library(here)
library(skimr) # install.packages("skimr")

```

## Read in our data

Our data is about lobsters

```{r}

lobsters <- read_excel(here("R-workshop-data", "lobsters2.xlsx"))

```

Explore our data

```{r}
# head() function shows the first 6 lines of data
head(lobsters)

# summary() to give a summary of each column
summary(lobsters)

## skimr::skim() gives additional summary statitics and is pipable ( %>% )
skim(lobsters)

# ggplot of lobsters
ggplot(data = lobsters, aes(x = year)) +
  geom_bar() +
  facet_wrap(~site)
```

## Our task:

How has the average size of lobsters changed over time across sites?

Pivot tables in R are group_by %>% summarize

```{r}

lobsters %>% # in my head I say "and then"
  group_by(year) %>% 
  summarize(countByYear = n())

```

What if we didn't group_by first?

```{r}

lobsters %>% 
  summarize(count_by_something = n())

## assing our summary to a new variable

lobstersSum <- lobsters %>% 
  summarize(count_by_something = n())

```

## group_by multiple variables

And summarize by multiple variables

We used mean() to calculate the mean size of lobsters by each site and year, and used na.rm = TRUE to calculate mean despite having missing values in our raw data.
```{r}

siteYearSummary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(countBySiteYear = n(),
            meanSizeMm = mean(size_mm, na.rm = TRUE),
            sdSizeMm = sd(size_mm, na.rm = TRUE))

siteYearSummary
```

## Activity

1. Calculate the median size_mm (?median)
2. Create and ggsave() a plot

```{r}

siteYearSummary <- lobsters %>% 
  group_by(site, year) %>% 
  summarize(countBySiteYear = n(),
            meanSizeMm = mean(size_mm, na.rm = TRUE),
            sdSizeMm = sd(size_mm, na.rm = TRUE),
            medianSizeMm = median(size_mm, na.rm = TRUE))

siteYearSummary

ggplot(data = siteYearSummary, aes(x = year, y = countBySiteYear, fill = site)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~site) +
  theme_minimal()

# I created a figures folder from the RStudio Files pane in order to be organized and for this code to run:
ggsave(here("figures/lobsterCountYear.png"), dpi = 180, width = 8, height = 7)

ggplot(data = siteYearSummary, aes(x = year, y = medianSizeMm, fill = site)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~site) +
  theme_minimal()

ggsave(here("figures/lobsterMedianYearFaceted.png"), dpi = 180, width = 8, height = 7)

ggplot(data = siteYearSummary, aes(x = year, y = medianSizeMm, color = site)) +
  geom_line() +
  theme_minimal()

ggsave(here("figures/lobsterMedianYear.png"), dpi = 180, width = 8, height = 7)

```

## dplyr::count() for counting by group
This is a "shortcut" function

```{r}

lobsters %>% 
  count(site, year)

## this is the same as

lobsters %>% 
  group_by(site, year) %>% 
  summarize(n = n()) %>% 
  ungroup()

```

# mutate() adds a new column
it will make a new variable

- Add a new column that is the size in meters
- Add a new column that is a repeated value
```{r}

head(lobsters)

lobsters %>% 
  mutate(sizeM = size_mm / 1000,
         millenia = 2000,
         observer = "Allison Horst")

```

## select() function to subset variables by columns

let's select
- date
- site
- sizeM

```{r}

lobstersSelected <- lobsters %>% 
  mutate(sizeM = size_mm / 1000) %>% 
  select(date, site, sizeM)

```




